---
title: "Test Exercise 3"
author: Jader Martins
date: \today
output:
  pdf_document:
    highlight: tango
    df_print: kable
  html_document:
    theme: united
    highlight: tango
---
```{r echo = F, include = F}
library(readxl)
library(tidyverse)
library(ggplot2)
library(knitr)
library(papeR)
library(lmtest)
data <- read_excel("TestExer-3-TaylorRule-round1.xlsx")
```



# (a) Use general-to-specific to come to a model. Start by regressing the federal funds rate on the other 7 variables and eliminate 1 variable at a time.

```{r}
reg.lm <- lm("INTRATE ~ INFL + PROD + UNEMPL + COMMPRI + PCE + PERSINC + HOUST", data)
pretty <- prettify(summary(reg.lm))
kable(pretty)
```
We remove PROD as it has the higher t value.
```{r}
reg.lm <- lm("INTRATE ~ INFL + UNEMPL + COMMPRI + PCE + PERSINC + HOUST", data)
pretty <- prettify(summary(reg.lm))
kable(pretty)
```

We remove UNEMPL as it has the higher t value.

```{r}
reg.lm <- lm("INTRATE ~ INFL + COMMPRI + PCE + PERSINC + HOUST", data)
pretty <- prettify(summary(reg.lm))
kable(pretty)
```

We remove COMMPRI as it has the higher t value.

```{r}
reg.lm <- lm("INTRATE ~ INFL + PCE + PERSINC + HOUST", data)
pretty <- prettify(summary(reg.lm))
kable(pretty)
summary(reg.lm)
```
Now all variables has high significance so we keep them.

# (b) Use specific-to-general to come to a model. Start by regressing the federal funds rate on only a constant and add 1 variable at a time. Is the model the same as in (a)?

```{r}
reg.lm <- lm("INTRATE ~ 1", data)
pretty <- prettify(summary(reg.lm))
kable(pretty)
```

```{r}
reg.lm <- lm("INTRATE ~ INFL", data)
pretty <- prettify(summary(reg.lm))
kable(pretty)
```

```{r}
reg.lm <- lm("INTRATE ~ INFL + PROD", data)
pretty <- prettify(summary(reg.lm))
kable(pretty)
```


```{r}
reg.lm <- lm("INTRATE ~ INFL + PROD + UNEMPL", data)
pretty <- prettify(summary(reg.lm))
kable(pretty)
```

```{r}
reg.lm <- lm("INTRATE ~ INFL + COMMPRI + UNEMPL", data)
pretty <- prettify(summary(reg.lm))
kable(pretty)
```


```{r}
reg.lm <- lm("INTRATE ~ INFL + PCE + UNEMPL", data)
pretty <- prettify(summary(reg.lm))
kable(pretty)
```


```{r}
reg.lm <- lm("INTRATE ~ INFL + PCE + PERSINC", data)
pretty <- prettify(summary(reg.lm))
kable(pretty)
```


```{r}
reg.lm <- lm("INTRATE ~ INFL + PCE + PERSINC + HOUST", data)
pretty <- prettify(summary(reg.lm))
kable(pretty)
summary(reg.lm)
```

As the final result we've got the same model as in (a).

# (c) Compare your model from (a) and the Taylor rule of equation (1). Consider $R^2$, AIC and BIC. Which of the models do you prefer?
```{r}
model1 <- lm("INTRATE ~ INFL + PCE + PERSINC + HOUST", data)
taylor <- lm("INTRATE ~ INFL + PROD", data)
AIC(model1, taylor)
BIC(model1, taylor)
summary(model1)$r.squared
summary(taylor)$r.squared
```
I prefer model1, as it explains better the variability in data.

# (d) Test the Taylor rule of equation (1) using the RESET test, Chow break and forecast test (with in both tests as break date January 1980) and a Jarque-Bera test. What do you conclude?
```{r}
resettest(data$INTRATE ~ data$INFL + data$PROD)
library(gap)
data$Year <- substr(data$OBS, start=1,stop=4)
grp <- data[data$Year < 1980, ]
x1 <- grp[, c("INFL", "PROD")]; y1 <- data.frame( INTRATE = grp["INTRATE"])
grp <- data[data$Year >= 1980, ]
x2 <- grp[, c("INFL", "PROD")]; y2 <- data.frame( INTRATE = grp["INTRATE"])
#chow.test
chow.test(y1, x1, y2, x2)
```
